---
title: "cm007 Exercises: Practice with `dplyr`"
output: 
  html_document:
    keep_md: true
    theme: paper
---

<!---The following chunk allows errors when knitting--->

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)
```

```{r}
#NOTES:
#summarize(new col sum name = function(col))

#within piped function there exists a new environment where the cols = vectors
#  ---> important for manipulation of hiplex material

#gapminder %>%  count(country) -> tells you how many within 

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tsibble))
```


This worksheet contains exercises aimed for practice with `dplyr`. 


1. (a) What's the minimum life expectancy for each continent and each year? (b) Add the corresponding country to the tibble, too. (c) Arrange by min life expectancy.

```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarize(min_life = min(lifeExp),
            country = country[lifeExp==min_life]) %>% 
  arrange(min_life)
```

2. Calculate the growth in population since the first year on record _for each country_ by rearranging the following lines, and filling in the `FILL_THIS_IN`. Here's another convenience function for you: `dplyr::first()`. 


```{r}
gapminder %>% 
  group_by(country) %>% 
  arrange(year) %>% 
  mutate(rel_growth = pop - first(pop)) %>% 
  DT::datatable()
```


3. Determine the country that experienced the sharpest 5-year drop in life expectancy, in each continent, sorted by the drop, by rearranging the following lines of code. Ensure there are no `NA`'s. Instead of using `lag()`, use the convenience function provided by the `tsibble` package, `tsibble::difference()`:

```{r}
 
 

 
 
 


 

```

```{r}
gapminder %>% 
  group_by(country) %>%
  arrange(year) %>% 
  mutate(inc_life_exp = difference(lifeExp)) %>%
  drop_na() %>%
  ungroup() %>%
  group_by(continent) %>%
  arrange(inc_life_exp) %>% 
  filter(inc_life_exp == min(inc_life_exp)) %>%
  knitr::kable()
```


